---
author: Lisa Hu
output:
  pdf_document:
    includes:
      before_body: title.sty
    toc: true
    toc_depth: 2
    number_sections: true
    latex_engine: xelatex
---

[//]: # (toc)

\newpage

```{r setup, include = TRUE, warning = FALSE, message = FALSE, results = "hide"}
#' Setup chunk
knitr::opts_chunk$set(cache = TRUE)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::read_chunk("EDA.R")
```

```{r libraries, warning = FALSE, message = FALSE}
#' Load all the packages
packages <- c("ggplot2", "tidyr", "dplyr", "readr", "tibble")
invisible(lapply(packages, library, character.only = T))
```

Dataset: [Urinary biomarkers for pancreatic cancer](https://www.kaggle.com/datasets/johnjdavisiv/urinary-biomarkers-for-pancreatic-cancer)

# Data description
## Reading the data
Running the following command:
```{r, eval = FALSE}
dataset <- read_excel("Data.xls")
```
Gave the following error:
```
filepath: <..>/Data.xls
libxls error: Unable to open file
```

Even though the filepath was correct, I tried different filenames and changing the method, but it still did not work. After scouring the internet, there was no clear solution to this. To avoid this error, open the xls-file in whichever program you have on hand for this file format. Depending on the program, `Export` (MS Excel) or `Save As` (LibreOffice Calc) the file to a CSV format (`Data.csv`).

From here, we can read the data and codebook:
```{r, read-data}
```

The information given in the codebook originates from the `Documentation.xls`. This file was given with the data file and can be found on the website.

## Manipulate and analyse the data
A lot of the rows contain empty strings instead of NA, so set them to NA first. Besides that, the columns `sample_id` and `benign_sample_diagnosis` in the dataset are not valuable for the analysis and are therefor dropped.

```{r, change-data}
```




