---
author: Lisa Hu
output:
  pdf_document:
    includes:
      before_body: title.sty
    keep_tex : true
    toc: true
    toc_depth: 2
    number_sections: true
    latex_engine: xelatex
---

[//]: # (toc)

\newpage

```{r setup, include = FALSE, warning = FALSE, message = FALSE, results = "hide"}
#' Setup chunk
knitr::opts_chunk$set(cache = TRUE)
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::read_chunk("EDA.R")
```

```{r libraries, warning = FALSE, message = FALSE, include = FALSE}
#' Load all the packages
packages <- c("ggplot2", "readr", "ggpubr", "pander", "missMethods", "ggbiplot")
invisible(lapply(packages, library, character.only = T))
```

# Data description
The data can be found on kaggle.com: [Urinary biomarkers for pancreatic cancer](https://www.kaggle.com/datasets/johnjdavisiv/urinary-biomarkers-for-pancreatic-cancer)
The files are saved as `Data.csv` and `Documentation.csv` for easier access.

The following packages were used:

+ ggplot2
+ tidyr
+ dplyr
+ readr

# Reading the data
We first want to create an insight of our data:
```{r read-data}
```

The information given in the codebook originates from the `Documentation.csv`. This file was given with the data file and can be found on the website.

\newpage
# Manipulate the data
A lot of the rows contain empty strings instead of NA, which has to be fixed first. Besides that, the columns `sample_id` , `patient_cohort`, `sample_origin`, and `benign_sample_diagnosis` in the dataset significant value for the analysis and are therefor dropped.

```{r change-data}
```

A summary of the data shows very high maximum values, but rather low medians. A log-transformation is applied to correct this. The missing values in the REG1A column will not be imputed since there is a lot of them and the imputation would only gravitate the data towards the imputation.

```{r log-trans}
```

The samples are then grouped by diagnosis for quicker access of the different samples. Table 5 shows the different amounts of samples per diagnosis and the amount of which are also blood samples. After the blood samples are seperated the column can be dropped.
\newpage
```{r demograph}
```

\newpage
# Analyse the data
```{r exp-box, fig.height = 8}
```

The outliers are not localized in a specific diagnosis group, but rather spread around everywhere.

\newpage

## Correlation matrix
```{r correlation}
```

The heatmap shows that there is not much correlation between creatinine and the other variables. The other outstanding one has to be the TFF1 biomarker, being the most correlated variable.

\newpage

## PCA
```{r pca}
```

The PCA shows that there is a clustering on the right upper side of the PDAC diagnosis. It is also very clear that creatinine has no correlation - as shown in the previous heatmap - but TFF1 does not seem as close to LYVE1 and REG1B as predicited. In fact, the latter two have a higher correlation with each other. Every point close to the origin have values close to the mean for all variables.
